Recursive Neural Network Part-of-Speech Tagger

### 2013-12-21 paper/paper.tex
Add a reference.
Add revisions plus the results table in the Analysis section.

To train the model, we first initialize $W_{combine}$, $b_{combine}$, $W_{classify}$ and $b_{classify}$ with small, uniformly distributed random values. We will discuss how to choose the range for these values in section \textbf{3.2.1 Weight Initialization} below. Then we perform a classification step in which we convert each word into its distributed word representation, pass each of these representations through the unrolled recursive neural network's inputs, and feed the values forward to the classification layers corresponding to each tag output.

@@

TODO: Describe future models.

@@

\subsubsection{Weight Initialization}

Early on, we randomly initialized our weight matrices and biases using Torch7's uniform tensor function,
\begin{lstlisting}
torch.rand(P, P+N)
torch.rand(P)
\end{lstlisting}
which draws from $\mathcal{U}[0, 1)$ however, we discovered guidelines for weight initialization from \hyperlink{http://deeplearning.net/tutorial/mlp.html\#mlp}{deeplearning.net/tutorial/mlp.html\#mlp}, borrowed from Glorot and Bengio (2010), and decided to explore how they impacted our training convergence.

@@

TODO: Describe analysis.

\subsubsection{Training Progress Diagram}
As we all know, there will be overfitting if we train too "hard" on training set. And if overfitting happens, the test accuracy, in our experiments is out-of-domain accuracy, will drop. Figure \ref{Adagrad_fig} shows this drop on both total accuracy and unknown word accuracy.

@@

TODO: Compare to baseline.

TODO: Compare to HMM.

\subsubsection{Error Analysis}

>>>>

To train the model, we first initialize $W_{combine}$, $b_{combine}$, $W_{classify}$ and $b_{classify}$ with small, uniformly distributed random values. We will discuss how to choose the range for these values in Section \ref{Weights} below. Then we perform a classification step in which we convert each word into its distributed word representation, pass each of these representations through the unrolled recursive neural network's inputs, and feed the values forward to the classification layers corresponding to each tag output.

@@

\paragraph{Bi-directional Recursive Neural Network}
From Section~\ref{sec:confm} Figure \ref{Pos_fig}, we find it is more likely to make mistakes on the end of sentences. It makes sense, because our basic RNN model uses only forward sequential information. So we think a bi-directional model will help, at least compared to the basic RNN. As for CVG model, which is simulating the HMM model considering the tags of whole sentence together, a bi-directional model may also help to kind of simulate CRF. There are two bi-direction models we are thinking:
\begin{itemize}
\item \textbf{Independently training and combining the probability: } Train two RNN models, and tag the sentence by combining the predicted probability from the two models' classifiers (after the softmax). 
\item \textbf{Jointly training: } Train two RNN models jointly, and concatenate two phrase representations and pass it to a shared classifier, which will in turn backward propagate the gradients to the two models so that the classifier part will get the forward and backward sequential information from two phrase representations. Figure~\ref{fig:bi-direc} shows the idea of this jointly training model.
\end{itemize}
\begin{figure}
\includegraphics[scale=0.5]{bi_direct.png}
\caption{Idea chart for jointly training bi-direction RNN model}\label{fig:bi-direc}
\end{figure}

\paragraph{Complex combination and classification model}
We are using one layer linear neural network on both combination and classification parts, because it is easy to train. In the beginning, we did try a two layer neural network, but it was too slow to train (about five times slower) and added another two hyper parameters to choose. However, because of the higher ability to capture more complex patterns, complex combination and classification may have better performance. Especially on classification part, we can use any machine learning classification model that can backward propagate gradients.

\paragraph{Update the lookup table}
In the beginning, we tried to update the lookup table together with other parameters. But the result was much worse than not updating it. But because deep learning researchers do update the feature learning layers (in our model, just lookup table), it is reasonable to try this. And instead of updating at the beginning of training, we can begin the updating after other parameters are well trained.

@@

\subsubsection{Weight Initialization}\label{Weights}

Early on, we randomly initialized our weight matrices and biases using Torch7's uniform tensor function,
\begin{lstlisting}
torch.rand(P, P+N)
torch.rand(P)
\end{lstlisting}
which draws from $\mathcal{U}[0, 1)$. However, we discovered guidelines for weight initialization from \hyperlink{http://deeplearning.net/tutorial/mlp.html\#mlp}{deeplearning.net/tutorial/mlp.html\#mlp}, borrowed from Glorot and Bengio (2010), and decided to explore how they impacted our training convergence.

@@

\subsubsection{Training Progress Diagram}
As we all know, there will be overfitting if we train too "hard" on training set. And if overfitting happens, the test accuracy (in our experiments, the out-of-domain accuracy) will drop. Figure \ref{Adagrad_fig} shows this drop on both total accuracy and unknown word accuracy.

@@

Below, we tally the results against the in-domain validation dataset and the out-of-domain validation dataset on all significantly performing recursive neural network models compared to the hidden-markov-model+TNT, plain hidden-markov-model and maximum likelihood tag models.

The AdaGrad variants started with a learning rate parameter at 0.1. The non-AdaGrad variants started with a learning rate of 0.0005. The RNN models trained on the entire training set (39815 sentences) were trained for 25 iterations. The RNN models trained on just 10,000 sentences were trained for 100 iterations.

Because we stripped all capitalization from our training and test data before converting words to their distributed vector representations, we noticed a high error rate due to conflating proper nouns with regular nouns, so to be fair to our model we evaluated the best model with no distinction between the two tags NNP and NN to see how it would fare had it properly considered capitalization as a feature.

\begin{center}
\resizebox{\columnwidth}{!}{%
\begin{tabular}{ l | ll | ll }
  Model & In & (In Unk) & Out & (Out Unk) \\
  \hline
  HMM+TNT & 95.88 & (87.13) & 93.67 & (84.61) \\
  RNN$_{39815}$ NN=NNP & 94.52 & (88.39) & 91.70 & (83.21) \\
  HMM & 95.52 & (63.95) & 91.23 & (53.00) \\
  AdaGrad RNN$_{39815}$ & 93.58 & (76.90) & 90.41 & (69.82) \\
  RNN$_{39815}$ & 93.47 & (77.06) & 90.28 & (70.43) \\
  AdaGrad RNN$_{10,000}$ & 93.02 & (79.17) & 90.12 & (73.43) \\
  RNN$_{10,000}$ & 92.72 & (77.51) & 89.54 & (69.60) \\
  Max Likelihood Tag & 91.31 & (42.74) & 87.01 & (53.31) \\
\end{tabular}%
}
\end{center}

\subsubsection{Error Analysis}  \label{sec:confm}
